---
title: "Data Wrangling Project"
author: "Lichen Yan"
date: "5/2/2021"
output: html_document
---
### dataset: https://www.kaggle.com/yannisp/uber-pickups-enriched
### Import packages
```{r}
library(ggplot2)
library(ggthemes)
library(lubridate)
library(dplyr)
library(tidyr)
library(scales)
library(knitr)
```

### Read the table and droping the columns that we don't need. 
```{r}
uber_nyc <- read.csv(file = "uber_nyc_enriched.csv")
head(uber_nyc)
uber_nyc <- select(uber_nyc, -c("spd", "vsb", "dewp", "slp", "pcp01", "pcp06", "pcp24"))
head(uber_nyc)
```

### Create servel new variables hour, day, dayofweek, month, year by the existing column pickup_dt. Then drop the Time and pickup_dt columns. 
```{r}
uber_nyc$Time <- format(as.POSIXct(uber_nyc$pickup_dt, format = "%Y-%m-%d %H:%M:%S"), format="%H:%M:%S")

uber_nyc$pickup_dt <- ymd_hms(uber_nyc$pickup_dt)
uber_nyc$day <- factor(day(uber_nyc$pickup_dt))
uber_nyc$month <- factor(month(uber_nyc$pickup_dt, label = TRUE))
uber_nyc$year <- factor(year(uber_nyc$pickup_dt))
uber_nyc$dayofweek <- factor(wday(uber_nyc$pickup_dt, label = TRUE))
uber_nyc$hour <- factor(hour(hms(uber_nyc$Time)))
uber_nyc$temp <- as.integer(uber_nyc$temp)
uber_nyc$sd <- as.integer(uber_nyc$sd)
uber_nyc <- select(uber_nyc, -c("Time", "pickup_dt"))
```

### Clean the data by drop all the NAs. 
```{r}
uber_nyc <- uber_nyc %>% drop_na()
head(uber_nyc)
```
### Creating a plot by hours in a day
```{r}
hour_data <- uber_nyc %>%
           group_by(hour) %>%
               dplyr::summarize(Total = sum(pickups))
ggplot(hour_data, aes(hour, Total)) + 
        geom_bar( stat = "identity", fill = "steelblue") +
           ggtitle("Pickups Every Hour") +
            theme(legend.position = "none") +
            scale_y_continuous(labels = comma)
```

### Creating a plot by hours and month
```{r}
month_hour <- uber_nyc %>%
          group_by(month, hour) %>%
             dplyr::summarize(Total = sum(pickups))

ggplot(month_hour, aes(hour, Total, fill = month)) + 
       geom_bar( stat = "identity", position='dodge') +
          ggtitle("Pickups by Hour and Month") +
           scale_y_continuous(labels = comma)
```

### Creating a plot by day
```{r}
day_group <- uber_nyc %>%
          group_by(day) %>%
             dplyr::summarize(Total = sum(pickups)) 

ggplot(day_group, aes(day, Total)) + 
        geom_bar( stat = "identity", fill = "steelblue") +
           ggtitle("Pickups Every Day") +
            theme(legend.position = "none") +
            scale_y_continuous(labels = comma)
```

### Creating a plot by dayofweek and month
```{r}
dayofweek_month_group <- uber_nyc %>%
         group_by(month, dayofweek) %>%
             dplyr::summarize(Total = sum(pickups))

ggplot(dayofweek_month_group, aes(month, Total, fill = dayofweek)) + 
        geom_bar( stat = "identity", position='dodge') +
           ggtitle("Pickups by Dayofweek and Month") +
            scale_y_continuous(labels = comma)
```

### Creating a plot by month
```{r}
month_group <- uber_nyc %>%
          group_by(month) %>%
             dplyr::summarize(Total = sum(pickups)) 

ggplot(month_group , aes(month, Total, fill = month)) + 
        geom_bar(stat = "identity", position='dodge') +
           ggtitle("Pickups by Month") +
            theme(legend.position = "none") +
            scale_y_continuous(labels = comma)
```

### Creating a plot by Temp
```{r}
temp_group <- uber_nyc %>%
          group_by(temp) %>%
             dplyr::summarize(Total = sum(pickups)/n()) 

ggplot(temp_group, aes(temp, Total)) + 
        geom_bar(stat = "identity", position='dodge') +
           ggtitle("Average pickups by Temp") +
            theme(legend.position = "none") +
            scale_y_continuous(labels = comma)
```
### Creating a plot by Snow depth
```{r}
temp_group <- uber_nyc %>%
          group_by(sd) %>%
             dplyr::summarize(Total = sum(pickups)/n()) 

ggplot(temp_group, aes(sd, Total)) + 
        geom_bar(stat = "identity", position='dodge') +
           ggtitle("Average pickups by sd") +
            theme(legend.position = "none") +
            scale_y_continuous(labels = comma)
```

### Creating a plot by county
```{r}
county_group <- uber_nyc %>%
          group_by(borough) %>%
             dplyr::summarize(Total = sum(pickups))

ggplot(county_group, aes(borough, Total)) + 
  geom_bar(stat = "identity", position='dodge') +
 scale_y_continuous(labels = comma) +
 ggtitle("Pickups by Borough")
```

### Creating a plot by county and dayofweek
```{r}
county_dayofweek_group <- uber_nyc %>%
          group_by(borough, dayofweek) %>%
             dplyr::summarize(Total = sum(pickups)) 

ggplot(county_dayofweek_group, aes(borough, Total, fill = dayofweek)) + 
 geom_bar(stat = "identity", position = "dodge") +
 scale_y_continuous(labels = comma) +
 ggtitle("Pickups by Borough and DayofWeek")
```

### Separate the data by holiday or not and giving the hypnosis that people are more likely to not take uber will on holiday.
```{r}
uber_nyc_Y <- subset(uber_nyc, hday %in% c("Y"))
uber_nyc_N <- subset(uber_nyc, hday %in% c("N"))

t.test(uber_nyc_N$pickups, uber_nyc_Y$pickups, alternative = "two.sided", conf.level = 0.95)
```

